A comprehensive analysis of natural sequence variation within the <i>Arabidopsis

### 2018-07-14 manuscript/02-methods.Rmd
-update gitignore for bookdown -update conditional table formatting -lots of revising -big rearrangement of repo structure

@@

$\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}$

@@

$\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
\qquad
\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
0 \quad \textrm{if} \quad N_{ik} = N_{jk}
\end{array}\right.$

@@

$\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
\qquad
\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}$

@@

so we can revert back to using Nei's original formula, modifying it slightly,
replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]

$\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}$

$\pi_{ij(k)} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad i \neq j \\
0 \quad \textrm{if} \quad i = j
\end{array}\right.$

@@

$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}$

@@

$\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})$

@@

$\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)$

@@

$\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)$

@@

>>>>

## Data Sources

### Variant Data
Variant data was queried from the 1001 genomes project website (http://1001genomes.org) via URL requests to an (undocumented) API service that provides subset files of the whole-genome VCF. The whole-genome VCF file can be found on the project’s website at http://1001genomes.org/data/GMI-MPI/releases/v3.1/

### Accession Information
Details on each of the 1135 accessions including CS stock numbers, and geographic location where the samples were collected were taken from the 1001 genomes website at http://1001genomes.org/accessions.html, via the download link at the bottom of the page. This data file has been embedded in the R package.

### Gene and transcript information
Information on the genes and transcripts including chromosomal coordinates, start and end location, and transcript length were taken from either the TAIR10 or Araport11 databases. The TAIR10 database, found at http://arabidopsis.org, was accessed via the biomart protocol, using the R package biomaRt. The araport11 database, which can also be found on the arabidopsis.org website (https://www.arabidopsis.org/download/index-auto.jsp?dir=%2Fdownload_files%2FGenes%2FAraport11_genome_release), has been embedded in the R package.

@@

\begin{equation}
\displaystyle \pi = \sum_{ij}x_{i}x_{j}\pi_{ij}
(\#eq:nei)
\end{equation}

@@

\begin{equation}
\displaystyle \pi = \frac{1}{L*n^2} \sum_{i=1}^n \sum_{j=1}^n \sum_{k=1}^L{\pi_{ijk}}
\qquad
\displaystyle \pi_{ijk} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad N_{ik} \neq N_{jk} \\
0 \quad \textrm{if} \quad N_{ik} = N_{jk}
\end{array}\right.
(\#eq:nei-gen)
\end{equation}

@@

\begin{equation}
\displaystyle \pi = \frac{1}{L} \sum_{k=1}^L{\pi_{k}}
\qquad
\pi_{k} = \frac{1}{n^2} \sum_{i=1}^n \sum_{j=1}^n {\pi_{ijk}}
(\#eq:nei-rearrange)
\end{equation}

@@

so we can revert back to using Nei's original formula \@ref(eq:nei), modifying it slightly,
replacing $x_i$ with $\frac{n_i}{n}$, $n_i$ being the number of sequences in the population of the form $N_i$ at location k: subscripts i and k are summed only over unique forms [need to clarify this]

\begin{equation}
\displaystyle \pi_k = \sum_{ij} \frac{n_i}{n}\frac{n_j}{n} \pi_{ij} = 
\frac{1}{n^2}\sum_{ij} n_i n_j \pi_{ij}\\
\pi_{ij(k)} = \left\{\begin{array}{l}
1 \quad \textrm{if} \quad i \neq j \\
0 \quad \textrm{if} \quad i = j
\end{array}\right.
(\#eq:nei-simp1)
\end{equation}

@@

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i n_{!i}
(\#eq:nei-simp2)
\end{equation}

@@

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} \sum_{i} n_i (n - n_{i})
(\#eq:nei-simp3)
\end{equation}

@@

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n \sum_{i} n_i - \sum_{i} n_{i}^2)
(\#eq:nei-simp4)
\end{equation}

@@

\begin{equation}
\displaystyle \pi_k = \frac{1}{n^2} (n^2 - \sum_{i} n_{i}^2)
(\#eq:nei-simp5)
\end{equation}

@@

## Software

The r1001genomes package has many software dependencies on other R packages, a few of the key bioinformatics packages used are listed below.

**biomaRt:** The biomaRt package was used for accessing the TAIR10 database on arabidopsis.org

**vcfR:** vcfR was used to read in the VCF files in a flat “tidy” format for easy manipulaiton

**BSgenome:** The BSgenome package BSgenome.Athaliana.TAIR.TAIR9 was used as the source for the complete DNA string of the reference genome (Col-0).

**DECIPHER:** DECIPHER was used to align nucleotide and amino acid sequences of homologous genes

**GenomicFeatures:** The GenomicFeatures package was used for handling sequence annotations.

**Biostrings:** Biostrings provides the underlying framework for the sequence manipulations used for generating and aligning sequences with BSgenome, Decipher, and GenomicFeatures
